%%%%%%%%%%%%%%%%%%%%%%acknow.tex%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% sample acknowledgment chapter
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%

\extrachap{Acknowledgments}

This report is the result of concerted effort by 50 experts in the field of Natural Hazards Engineering. The editors are grateful for every contribution, feedback, and suggestion received since the first edition of the report was published in 2019. Special thanks are due to the twelve corresponding authors who organized and distilled content from various sources into 22 chapters. Finally, we would like to express our gratitude to Claire M. Johnson for editing and Claudio Perez for his invaluable help with formatting the document.


The identification of certain commercial systems and research tools in this report does not imply recommendation or endorsement by the National Science Foundation, the Computational Modeling and Simulation Center (SimCenter), or the academic institutions of contributing authors. Nor does such identification imply that those products are necessarily the best available for the task. 


This material is based upon work supported by the National Science Foundation under Grant No. 1612843. Any opinions, findings, and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the National Science Foundation or the Regents of the University of California.

