%%%%%%%%%%%%%%%%%%%%% appendix.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{List of Software Tools}

The tables in this appendix were prepared to provide an overview of a few important features of the tools mentioned in each chapter. The presented information is meant to help researchers compare the available tools and see which one fits their needs. 


The type of license and supported platforms are based on the available descriptions and data on each software's website. An empty license means that the developers have not provided that information; it does not necessarily mean that the software is freeware. Even though applications with an open source code could often be compiled on all three platforms considered here, the information in the tables corresponds to their availability for typical usage. That is, if there are pre-built binaries available and typical users would download those binaries, then the information in the tables refers to the platforms supported by those binaries.


Ticks in the DesignSafe column identify applications that are either available in the Workspace on the DesignSafe website or they can take advantage of DesignSafe resources by establishing a secure remote connection to the service. 


The identification of certain commercial systems and research tools in the tables does not imply recommendation or endorsement by the SimCenter or the academic institutions of contributing authors. Nor does such identification imply that those products are necessarily the best available for the task.


The abbreviations used in the tables are described below.

\input{keys.tex}

\input{tables.tex}
